---
title: "NeuroImage:Reports - Aggregating Data"
author: "<h3>by Michael Demidenko</h3>"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: yes
    number_sections: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    code_folding: hide
    self_contained: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
tags: []
subtitle: <h2><u>Extension and Multiverse assessment </u></h2>
---
# Project Description
This is an extension and multiverse analysis of a publication in Development and Psychopathology ([2020](www.doi.org/10.1017/S0954579419000580));

*Pubertal development mediates the association between family environment and brain structure and function in childhood.* 
> By Sandra Thijssen, Paul Collins and Monica Luciana.

This project is co-authored by:
Michael I. Demidenko, Dominic P. Kelly, Felicia Hardi, Ka I. Ip, Sujin Lee, Sunghyun Hong, Sandra Thijssen, Monica Luciana, Hannah Becker, & Daniel P Keating

> This Rmarkdown file contains the code for aggregating the ABCD variables


*Note: the option for this rmarkdown is `code_folding: hide`, so the code is hidden. Please select `Code` --> `Show All Code` to unhide all code blocks. Otherwise, in each section you can simply click the `Code` button to show the respective code chunk.*

Useful R/Mplus links:

* Base R [cheatsheets](https://www.povertyactionlab.org/sites/default/files/r-cheat-sheet.pdf) 
* Using [Tabsets](https://bookdown.org/yihui/rmarkdown-cookbook/html-tabs.html)
* Statistical Analyses in R from [Russel Poldrack](https://statsthinking21.github.io/statsthinking21-R-site/)
* Instructional videos from [Karandeep Singh](https://www.youtube.com/channel/UC7eRNr1Pprls2aU5WWPz15Q/videos)

***
***

# Preparing packages {.tabset}

In order to perform the replication and extension of the original [work](www.doi.org/10.1017/S0954579419000580), it is necessary to identify all of the variables and prepackages these variables from the downloaded NDA dataset. Before that, we install and load the necessary packages and functions, then pull and aggregate the files and

After this step is complete, you can work on the second Rmarkdown file.
The associated variables and data dictionary labels are covered below.

## Install packages 

Here I include a list of packages that are checked and automatically installed that will be required to run this script from start-to-finish.

```{r message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse)
```

## Load packages

At each instance, load packages.
```{r message=FALSE, warning=FALSE}
# Load packages
library(tidyverse) 
```

# Prepare data {.tabset}

## Aggregate data 

Here we use code from `Mike Angstadt`, available on [github](https://github.com/mangstad/Misc_utils/blob/master/abcd_functions.R) which uses an R [function](https://www.tutorialspoint.com/r/r_functions.htm) to aggregate data across text files from the downloaded ABCD [data repository](https://nda.nih.gov/abcd). 
`read.abcd` function readings in the files, tab separated, skipping 1 row and not including descriptives. It pulls remaining data and columns into table. Then, `multi.merge` takes in different dataframes that consistent of important text files and joins then `by` specified variables (we focus on subject key and event name, as these are consistent (or should be) across the .txt files).

```{r message=FALSE, warning=FALSE}
read.abcd = function(file, sep="\t",skip=1,cols=NULL,descriptions=FALSE) {
  headers = names(read.table(file,sep=sep,header=T,stringsAsFactors=F)[-1,])
  if (descriptions) {
    descrip = names(read.table(file,sep=sep,header=T,stringsAsFactors=F,skip=1))
  }
  data = read.table(file,sep=sep,header=T,stringsAsFactors=F,skip=skip)
  names(data) = headers
  if (!is.null(cols)) {
    data = subset(data,select=cols)
  }
  if (descriptions) {
    list(data=data,descrip=descrip)
  } else {
    data
  }
}

multi.merge = function(...,by=NULL) {
  Reduce(function(x,y) merge(x,y, all=TRUE,by=by), list(...))
}

```

## Compile ABCD data

Using the above functions, we cook up the dataset. Specifically, we set `mergecols` to the values that we want to merge by. We us [c()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/c) to concatenate values into a list. As mentioned, our list consists of subjectkey and eventname. Then, using the `read.abcd` function defined above, we read in the file, and using the `%>%` or [pipe](https://style.tidyverse.org/pipes.html) function in the conjunction with [select()](https://dplyr.tidyverse.org/reference/select.html) we subset the values that we want to pull from that data. For example, for *read.abcd("ABCDStudyNDA_1182451/pdem02.txt")* we subset the subjectkey and eventname that will be used for `by` in column merge, and then our variables of interest. The variables we used are defined and labeled on the 

The variables identified in the manuscript we cross-reference for the associated text file name on the NDA NIMH [data dictionary ABCD release 3.0](https://nda.nih.gov/data_dictionary.html?submission=ALL&source=ABCD%2BRelease%2B3.0). The name for the .txt file is associated with the `short name` on the NIMH data archive. Within each variable we then pull the associated variables. The list of these are below:

  
**Demographics**:

*	*General*: age, sex, race/ethnicity, family ID and child sibling status [acspsw03](https://nda.nih.gov/data_structure.html?short_name=acspsw03)
* *Factor*:
  + Demographic Income, Education, biological parents, and who guardian during visit is in [pdem02](https://nda.nih.gov/data_structure.html?short_name=pdem02)
  + Planned pregnancy [devhx_6_p](https://nda.nih.gov/data_structure.html?short_name=dhx01)
  + total problems score, Raw, t score ASEBA[abcd_asrs01](https://nda.nih.gov/data_structure.html?short_name=abcd_asrs01) 
* Site ID [abcd_lt01](abcd_lt01)
* Biological parent [pdem02](https://nda.nih.gov/data_structure.html?short_name=pdem02)

**Child-self reported Questions**

* Parental monitory survey [pmq01](https://nda.nih.gov/data_structure.html?short_name=pmq01)
* Maternal Acceptance CRPBI [crpbi01](https://nda.nih.gov/data_structure.html?short_name=crpbi01)
* Family Environment Scale (Phenx) Youth [abcd_fes01](https://nda.nih.gov/data_structure.html?short_name=abcd_fes01)

**Parent Reported Information**

* Family Environment Scale (Phenx) Parent [fes02](https://nda.nih.gov/data_structure.html?short_name=fes02)
* KSADS item measuring parent/child relationship [dibf01](https://nda.nih.gov/data_structure.html?short_name=dibf01)

**Puberty Measures**

* Tabulated PDS Puberty value for Female [abcd_ssphp01](https://nda.nih.gov/data_structure.html?short_name=abcd_ssphp01)
* Tabulated PDS Puberty value for Female [abcd_ssphy01](https://nda.nih.gov/data_structure.html?short_name=abcd_ssphy01)
* Individual PDS Puberty Items for Youth Report [abcd_ypdms01](https://nda.nih.gov/data_structure.html?short_name=abcd_ypdms01)
* Individual PDS Puberty Items for Parent Report [abcd_ppdms01](https://nda.nih.gov/data_structure.html?short_name=abcd_ppdms01)

**Brain Measures**

*Gray Matter*

* Amygdala volume, L/R [abcd_smrip201](https://nda.nih.gov/data_structure.html?short_name=abcd_smrip201)
* Anterior Cingulate Cortex Thickness & volume, L/R [abcd_mrisdp101](https://nda.nih.gov/data_structure.html?short_name=abcd_mrisdp101)
* Total Cortical volume (regressed out) [abcd_mrisdp101](https://nda.nih.gov/data_structure.html?short_name=abcd_mrisdp101)
*	QC T1 Motion, WM Underestimation, Inhomogeneity score, pial overestimation score, artifact score [freesqc01](https://nda.nih.gov/data_structure.html?short_name=freesqc01)

*White Matter*

* Rostral/caudal Anterior Cingulate FA, L/R [abcd_dti_p101](https://nda.nih.gov/data_structure.html?short_name=abcd_dti_p101)
*	All WM Tract volume (regressed out) [abcd_dti_p101](https://nda.nih.gov/data_structure.html?short_name=abcd_dti_p101)
*	QC WTI Framewise Displacement, motion [abcd_dti_p101](https://nda.nih.gov/data_structure.html?short_name=abcd_dti_p101)
    
*Resting state*

*	L/R Amygdala – Cingulo-Opercular Network connectivity [mrirscor02](https://nda.nih.gov/data_structure.html?short_name=mrirscor02)
*	Function Connectivity Framewise displacement, motion [mrirscor02](https://nda.nih.gov/data_structure.html?short_name=mrirscor02)
*	Resting state fieldmap wrong locus [mriqcrp102](https://nda.nih.gov/data_structure.html?short_name=mriqcrp102)


```{r echo=TRUE}
#merging by
mergecols = c("subjectkey","eventname")


# Demographics + parental education + parental psychopathology, relationship status, pregnancy

par_inc_edu = read.abcd("ABCDStudyNDA_1182451/pdem02.txt") %>% 
  select(subjectkey,eventname,
         demo_comb_income_v2, 
         demo_prnt_ed_v2,
         demo_prim, # you are the: 1 = Childs Biological Mother 2 = Childs Biological Father 3 = Adoptive Parent ; 4 = Childs  5 = Other Otro,
         demo_prnt_marital_v2, # are you now married, widowed, divorced, separated, living with partner
         demo_prnt_prtnr_bio # is your partner the child's biological parent? (1/0)
         )


planned_pre = read.delim("ABCDStudyNDA_1182451/dhx01.txt",sep = "\t", header = TRUE)[-1,] %>% rename(eventname = visit) %>% 
  select(subjectkey, eventname,
         devhx_6_p) # used diff code due to compiling error

par_psychopath = read.abcd("ABCDStudyNDA_1182451/abcd_asrs01.txt") %>% 
  select(subjectkey,eventname,
         asr_scr_totprob_total, # total problems score
         asr_scr_totprob_r, # raw score ASEBA
         asr_scr_totprob_t #t score in relation to norms for each gender at 18-35 and 36-59, based on national probability samples
  )





# ***Child-reported questionnaires*** included an abbreviated version of the 

#  Parental Monitoring Survey (Chilcoat & Anthony, 1996). 
par_monitoring = read.abcd("ABCDStudyNDA_1182451/pmq01.txt") %>% 
  select(subjectkey,eventname,
         parent_monitor_q1_y:parent_monitor_q5_y
         )

# maternal acceptance scale of the Child Report of Parent Behavior Inventory (Schaefer, 1965) -- there is an item for a *second caregiver* too
child_accept = read.abcd("ABCDStudyNDA_1182451/crpbi01.txt") %>% 
  select(subjectkey,eventname,
         crpbi_parent1_y:crpbi_parent5_y
         )

child_conflict =  read.abcd("ABCDStudyNDA_1182451/abcd_fes01.txt") %>% 
  select(subjectkey,eventname,
         fes_youth_q1:fes_youth_q9
         )
# ***parent-reported information*** included an abbreviated version of the 

# the conflict scale from the Family Environment Scale (FES; Moos & Moos, 1976)
family_conflict = read.abcd("ABCDStudyNDA_1182451/fes02.txt") %>% 
  select(subjectkey,eventname,
         fam_enviro1_p:fam_enviro9r_p
         )
# Complimented by one background item from the Kiddie Schedule for Affective Disorders and Schizophrenia (Kaufman et al., 1997) measuring the relationship between the parent and the child (“in general, how do you and your child get along?” using a 3-point Likert scale ranging from very well to a lot of conflict).
kiddie_parchildrela = read.abcd("ABCDStudyNDA_1182451/dibf01.txt") %>% 
  select(subjectkey,eventname,
         kbi_p_conflict
         )



# select puberty
final_p_pubert = read.abcd("ABCDStudyNDA_1182451/abcd_ssphp01.txt") %>%
  select(subjectkey,eventname,
         pds_p_ss_female_category_2,
         pds_p_ss_male_category_2)

final_y_pubert = read.abcd("ABCDStudyNDA_1182451/abcd_ssphy01.txt") %>%
  select(subjectkey,eventname,
         pds_y_ss_female_category_2,
         pds_y_ss_male_cat_2)

youth_puberty = read.abcd("ABCDStudyNDA_1182451/abcd_ypdms01.txt") %>%
  select(subjectkey,eventname,
         y_pds_ht = pds_ht2_y,
         y_pds_skin = pds_skin2_y,
        y_pds_bdhair =  pds_bdyhair_y,
         y_pds_brstgrw = pds_f4_2_y, # breasts growth
         y_pds_menstrual = pds_f5_y, # menstruation
         y_pds_dpvoic = pds_m4_y, # deep voice
         y_pds_fcHair = pds_m5_y #hair face
         )

parent_puberty = read.abcd("ABCDStudyNDA_1182451/abcd_ppdms01.txt") %>%
  select(subjectkey,eventname,
         p_pds_ht = pds_1_p,#height
         p_pds_bdhair = pds_2_p,#body hair
         p_pds_skin = pds_3_p,#changes skin
         p_pds_fcHair = pds_m5_p, # hair face
         p_pds_dpvoic = pds_m4_p, # deep voice
         p_pds_brstgrw = pds_f4_p, # breast growth
         p_pds_menstrual = pds_f5b_p #menstruate
         )

# rel_fam & race
demograph = read.abcd("ABCDStudyNDA_1182451/acspsw03.txt") %>%
  select(subjectkey,eventname,
         interview_age,
         sex,# M = Male; F = Female; O=Other; NR = Not reported
         race_ethnicity, #1 = White; 2 = Black; 3 = Hispanic; 4 = Asian; 5 = Other
         rel_family_id,
         rel_relationship # 0 = single; 1 = sibling; 2 = twin; 3 = triplet
         )


# site id
site = read.abcd("ABCDStudyNDA_1182451/abcd_lt01.txt") %>% 
  select(subjectkey,eventname,
         interview_age,
         interview_date,
         abcd_site = site_id_l)

## BRAIN ###

# GM structural
sub_struct = read.abcd("ABCDStudyNDA_1182451/abcd_smrip201.txt") %>% 
  select(subjectkey,eventname,
         L_Amyg_vol = smri_vol_scs_amygdalalh, 
         R_Amyg_vol = smri_vol_scs_amygdalarh
         )

cort_struct = read.abcd("ABCDStudyNDA_1182451/abcd_mrisdp101.txt") %>% 
  select(subjectkey,eventname,
         L_ACC_thic = mrisdp_6, R_ACC_thic = mrisdp_80,
         L_ACC_area = mrisdp_308, R_ACC_area = mrisdp_382,
         L_Cort_vol = mrisdp_602, R_Cort_vol = mrisdp_603)# fixed

QC_struct = read.abcd("ABCDStudyNDA_1182451/freesqc01.txt") %>% 
  select(subjectkey,eventname,
         QC_T1_motion = fsqc_qu_motion, QC_T1_inhomog = fsqc_qu_inhomogeneity,
         QC_T1_wmunder = fsqc_qu_wmunder, QC_T1_pial = fsqc_qu_wmunder, QC_T1_artifact = fsqc_qu_artifact)

# WM structural
dti_struct = read.abcd("ABCDStudyNDA_1182451/abcd_dti_p101.txt") %>% 
  select(subjectkey,eventname,
         L_Rost_ACC = dmri_dtifawm_cdsn_rncntelh, 
         R_Rost_ACC = dmri_dtifawm_cdsn_rncnterh,
         L_Caud_ACC = dmri_dtifawm_cdsn_cdancntelh,
         R_Caud_ACC = dmri_dtifawm_cdsn_cdancnterh,
         QC_dti_FDmot = dmri_dti_meanmotion,
         DTI_vol_alltract = dmri_dtivol_fiberat_allfibers
         )

# Resting state
rsfmri = read.abcd("ABCDStudyNDA_1182451/mrirscor02.txt") %>% 
  select(subjectkey,eventname,
         Lamyg_CingOperc = rsfmri_cor_ngd_cerc_scs_agrh,
         Ramyg_CingOperc = rsfmri_cor_ngd_cerc_scs_aglh,
         FDmot_FC = rsfmri_cor_ngd_scs_meanmn
         )

# Fieldmaps
rsfm_field = read.abcd("ABCDStudyNDA_1182451/mriqcrp102.txt") %>% 
  select(subjectkey,eventname,
         rsfmri_fieldmap_0 = iqc_rsfmri_1_fm_wronglocus
         )


# Merge all above by site/id
abcd = multi.merge(family_conflict, kiddie_parchildrela, 
                   par_monitoring, child_accept,  child_conflict,
                   par_inc_edu, par_psychopath, planned_pre,
                   final_p_pubert,final_y_pubert, 
                   youth_puberty,parent_puberty, demograph, site, 
                   sub_struct, cort_struct,QC_struct,dti_struct,rsfm_field,rsfmri,
                   by=mergecols)

```


Now we use the [tidyverse](https://www.rstudio.com/resources/cheatsheets/) to filter our data by baseline data in `eventname`. The release 3.0 ABCD data consists of Baseline, 6 month and 1 year data. Given that the original manuscript used the first half of Baseline and we want to use the *full baseline data* we restrict our data to baseline.
Also, in some cases there are columns that a repeated in the .txt files which may result in repeat cases. To avoid this, we use the [duplicated](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/duplicated) command to restrict our data to `unique` subjects.

```{r message=FALSE, warning=FALSE}
abcd = abcd %>% filter(eventname == "baseline_year_1_arm_1") %>% 
  mutate(event_name = case_when(eventname == "baseline_year_1_arm_1" ~ "Baseline"))

abcd = abcd[!duplicated(abcd),] # created redundant rows, remove
```

## Export aggregated data to local 

Export the aggregated `abcd` data into a .csv file using [write.csv](https://www.rdocumentation.org/packages/AlphaPart/versions/0.8.4/topics/write.csv). This cooked up data will surve as the starting point for the next phase.

```{r}
# write file
write.csv(abcd, "ABCD_cooked_2021_11_15.csv", row.names = FALSE)
```

